---
title: "SSD project"
author: "Anthony"
date: "01/14/2022"
output: html_document
---

#clear the memory
```{r}
rm(list=ls())
```

#library
```{r}
library(ggplot2)
library(tidyr)
library(corrplot)
library(dplyr)
library(date)
library(lubridate)
library(tidyverse)
library(knitr)
library(ggpubr)
```

#set up directory
```{r}
setwd("~")
list.files()
```

#load data
```{r}
ssd<-read.csv("national_ssd.csv")
pulse<-read.csv("pulse_revise.csv")
nssp<-read.csv("NSSP.csv")
```

#clean pulse dataset
```{r}
anxiety<-pulse %>% filter(Indicator=="Symptoms of Anxiety Disorder") %>% select(Time.Period,Value)
depression<-pulse %>% filter(Indicator=="Symptoms of Depressive Disorder") %>% select(Time.Period,Value)
pres<-pulse %>% filter(Indicator=="Took Prescription Medication for Mental Health, Last 4 Weeks" | Indicator=="Received Counseling or Therapy, Last 4 Weeks" | Indicator=="Needed Counseling or Therapy But Did Not Get It, Last 4 Weeks") %>% select(Indicator,Time.Period,Value)

anxiety<-anxiety[1:27,]
depression<-depression[1:27,]
anxiety<-anxiety %>% rename(week=Time.Period)
depression<-depression %>% rename(week=Time.Period)
pres<-pres %>% rename(week=Time.Period)
pres$Indicator<-ifelse(pres$Indicator=="Took Prescription Medication for Mental Health, Last 4 Weeks","Meds",ifelse(pres$Indicator=="Received Counseling or Therapy, Last 4 Weeks","Therapy","Unmet"))
pres<-spread(pres,Indicator,Value)

#apply scale factor 1/100
anxiety$Value<-anxiety$Value/100
depression$Value<-depression$Value/100
pres$Meds<-pres$Meds/100
pres$Therapy<-pres$Therapy/100
pres$Unmet<-pres$Unmet/100
  
```

#SAME WEEK
##match with Google search(same week)
```{r}
anxiety$time<-ifelse(anxiety$week<13,anxiety$week+16,
                          ifelse(anxiety$week>21,anxiety$week*2-43,anxiety$week*2+8))
depression$time<-ifelse(depression$week<13,depression$week+16,
                             ifelse(depression$week>21,depression$week*2-43,depression$week*2+8))
pres$time<-ifelse(pres$week<13,pres$week+16,
                       ifelse(pres$week>21,pres$week*2-43,pres$week*2+8))

#subset data ssd
symptom<-c("symptom.Anxiety","symptom.Depression","symptom.Motion.sickness",
           "symptom.Suicidal.ideation","date")
ssd_new<-ssd[symptom]
ssd_new$time<-mdy(ssd_new$date)
ssd_new$week<-floor_date(ssd_new$time,"week")

#get week number from date
ssd_new$time_new<-strftime(ssd_new$week,format="%V")
ssd_new$time_new<-as.numeric(ssd_new$time_new)
#cut off date
ssd_new<-ssd_new[-53,]
ssd_new<-ssd_new[-66,]
```

#deal with nssp data
```{r}
ed<-c("Total.Number.of.ED.Visits.by.Week","Disaster.Related.Mental.Health.Age.10..Count",
           "Suicide.Attempt.Age.10..Count","All.Drug.Overdose.Count.Age.10.",
           "Opioid.Overdose.Count.Age.10.","Intimate.Partner.Violence.Age.18..Count","Total.ED.Visits.by.Week.Age.0.17","Suspected.Child.Abuse.and.Neglect.Age.0.17.Count","Week.Beginning.Date")
nssp_new<-nssp[ed]

nssp_new$time<-dmy(nssp_new$Week.Beginning.Date)
nssp_new$week<-floor_date(nssp_new$time,"week")
nssp_plot<-nssp_new %>% select(Disaster.Related.Mental.Health.Age.10..Count,Suicide.Attempt.Age.10..Count,time)

```

#deal with pulse data
```{r}
ssd_pulse<-ssd_new[-1:-13,]
ssd_pulse<-ssd_pulse %>% filter(year(time) %in% c(2020,2021))
pulse_anxiety<-merge(ssd_pulse,anxiety,by.x = "time_new", by.y = "time",all.x = FALSE,all.y = TRUE)
pulse_depression<-merge(ssd_pulse,depression,by.x = "time_new", by.y = "time",all.x = FALSE,all.y = TRUE)
pulse_pres<-merge(ssd_pulse,pres,by.x = "time_new", by.y = "time",all.x = FALSE,all.y = TRUE)
pulse_anxiety<-pulse_anxiety[,-1:-7]
pulse_depression<-pulse_depression[,-1:-7]
pulse_pres<-pulse_pres[,-1:-7]
pulse<-merge(pulse_anxiety,pulse_depression,by.x = "week.x",by.y = "week.x")
pulse<-merge(pulse,pulse_pres,by.x = "week.x",by.y = "week.x",all.x = TRUE)
pulse<-pulse[,-2]
pulse<-pulse[,-3]
pulse<-pulse[,-4]
```


#prepare data for figures 
```{r}
#merge ssd_new, pulse, nssp_plot to ensure the same x axis
m_data<-merge(ssd_new,nssp_plot, by.x = "week", by.y = "time", all.x = TRUE)
m_data<-merge(m_data,pulse, by.x = "week", by.y = "week.x", all.x = TRUE)

#reconstruct ssd_new, pulse, nssp_plot
ssd_new<-m_data[,c(1:5)]
pulse<-m_data[,c(1,11:15)]
nssp_plot<-m_data[,c(1,9,10)]

```

#prepare data for ploting figure 1 
```{r}
m_data$Year<-year(m_data$week)
m_data_plot<-m_data %>% filter(Year %in% c(2018,2019,2020)) %>%
  select(week,symptom.Anxiety,symptom.Depression,symptom.Suicidal.ideation,Disaster.Related.Mental.Health.Age.10..Count,Suicide.Attempt.Age.10..Count,Year)
m_data_ssd<-m_data_plot %>% select(week, symptom.Anxiety,symptom.Depression,symptom.Suicidal.ideation,Year)
m_data_nssp<-m_data_plot %>% select(week, Disaster.Related.Mental.Health.Age.10..Count,Suicide.Attempt.Age.10..Count,Year)

#number the week
m_data_ssd$Year<-factor(m_data_ssd$Year)
m_data_ssd$date<-as.Date(strftime(m_data_ssd$week,format = "%m-%d"),format="%m-%d")
m_data_nssp$Year<-factor(m_data_nssp$Year)
m_data_nssp$date<-as.Date(strftime(m_data_nssp$week,format = "%m-%d"),format="%m-%d")

```


#fit GLM model to test significant decline for holiday
```{r}
df<-m_data_ssd %>% mutate(thx=case_when(week %in% as.Date(c("2018-11-18",
                                                            "2019-11-24",
                                                            "2020-11-22"))~1, TRUE~0),
                          xmas=case_when(week %in% as.Date(c("2018-12-23",
                                                             "2019-12-22",
                                                             "2020-12-20"))~1, TRUE~0),
                          covid=case_when(week %in% as.Date(c("2020-03-01","2020-03-08",
                                                              "2020-03-15","2020-03-22"))~1, TRUE~0),
                          month=month(week),
                          season=case_when(month %in% c(2,3,4)~'spring',
                                           month %in% c(5,6,7)~'summer',
                                           month %in% c(8,9,10)~'fall',
                                           TRUE~'winter'))

model_an<-glm(symptom.Anxiety ~ Year + season + thx + xmas + covid, data = df, family = gaussian(link = 'log')  )
summary(model_an)
1-exp(confint(model_an))

model_de<-glm(symptom.Depression ~ Year + season + thx + xmas + covid, data = df, family = gaussian(link = 'log')  )
summary(model_de)
1-exp(confint(model_de))

model_su<-glm(symptom.Suicidal.ideation ~ Year + season + thx + xmas + covid, data = df, family = gaussian(link = 'log')  )
summary(model_su)
1-exp(confint(model_su))

```
#plot combined figure1
```{r}
m_data_ssd_long<-gather(m_data_ssd,symptom,value,2:4,factor_key = TRUE)
ggplot(m_data_ssd_long,aes(x=date,y=value))+aes(shape=symptom)+
  geom_line(aes(color=Year),size=1)+geom_vline(xintercept=as.Date("2021-03-07"), linetype="dashed",size=1)+
  xlab("Time")+ylab("Normalized Search Volume")+
  scale_y_continuous(limits=c(0,12),
                     breaks = c(0,1.5,3.0,4.5,6.0,7.5,9.0,10.5,12.0))+
  scale_x_date(date_labels = "%b",breaks = as.Date(c("2021-01-07","2021-02-07","2021-03-07","2021-04-07","2021-05-07","2021-06-07","2021-07-07","2021-08-07","2021-09-07","2021-10-07","2021-11-07","2021-12-07")))+theme_bw()+
  theme(axis.text.x = element_text(color = "grey20", size =18),
        axis.text.y = element_text(color = "grey20", size = 18),  
        axis.title.x = element_text(color = "grey20", size = 20),
        axis.title.y = element_text(color = "grey20", size = 20),
        axis.line = element_line(color="black"),
        legend.text=element_text(size=18),
        legend.title = element_text(size=20),
        panel.border = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank())
ggsave("figure1_new.eps",height=12,width = 15)

```

#interpolate pulse data with Excel 
```{r}
#interpolate pulse data
## Step 1: construct and save dataset for interpolation 
#pulse_new<-pulse %>% filter(year(week) %in% c(2020,2021))
#write.csv(pulse_new,"pulse_interpolate.csv")
## Step 2: do interpolation with EXCEL
## Step 3: mute the step 1 codes and load interpolated pulse data
pulse_new<-read.csv("pulse_interpolate.csv")
pulse_new<-pulse_new[,-1]
pulse_new$week<-mdy(pulse_new$week)
```

#plot figure 2, percentage change between 2019 and 2020
```{r}
#interpolate percentage change using Excel
## Step 1: save dataset for manipulation
#write.csv(m_data_plot,"percent_change.csv",row.names = FALSE)
## Step 2: mute the step 1 codes and load percentage change data
per_change<-read.csv("percent_change.csv")
per_change<-per_change %>% select(week,anxiety,depression,suicidal,mental,attemp)
per_change$week<-mdy(per_change$week)
per_change$date<-as.Date(strftime(per_change$week,format = "%m-%d"),format="%m-%d")
pulse_new<-pulse_new %>% filter(year(week)<2021)

r<-ggplot(per_change,aes(x=date,group = 1))
r<-r+geom_line(aes(y=anxiety,color="Anxiety"),size=1)+geom_line(aes(y=depression,color="Depression"),size=1)+
  geom_line(aes(y=suicidal,color="Suicidal Ideation"),size=1)+geom_vline(xintercept=as.Date("2021-03-07"), linetype="dashed",size=1)+labs(y="Percent Change \n in Search ", x="Time", color="SSD Term")+scale_y_continuous(labels = scales::percent)+scale_x_date(date_labels = "%b",breaks = as.Date(c("2021-01-07","2021-02-07","2021-03-07","2021-04-07","2021-05-07","2021-06-07","2021-07-07","2021-08-07","2021-09-07","2021-10-07","2021-11-07","2021-12-07")))+theme_bw()+
  theme(axis.text.x = element_text(color = "grey20", size =16),
        axis.text.y = element_text(color = "grey20", size = 16),  
        axis.title.x = element_text(color = "grey20", size = 16),
        axis.title.y = element_text(color = "black", size = 18, face = "bold"),
        legend.text=element_text(size=16),
        legend.title = element_text(size=18),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank())

s<-ggplot(per_change,aes(x=date,group = 1))
s<-s+geom_line(aes(y=mental,color="Disaster Related \nMental Health"),size=1)+geom_line(aes(y=attemp,color="Suicidal Attempt"),size=1)+geom_vline(xintercept=as.Date("2021-03-07"), linetype="dashed",size=1)+
  labs(y="Percent Change \n in NSSP ED Visits", x="Time", color="Cause of ED Visit")+
  scale_y_continuous(labels = scales::percent)+scale_x_date(date_labels = "%b",breaks = as.Date(c("2021-01-07","2021-02-07","2021-03-07","2021-04-07","2021-05-07","2021-06-07","2021-07-07","2021-08-07","2021-09-07","2021-10-07","2021-11-07","2021-12-07")))+theme_bw()+
  theme(axis.text.x = element_text(color = "grey20", size =16),
        axis.text.y = element_text(color = "grey20", size = 16),  
        axis.title.x = element_text(color = "grey20", size = 16),
        axis.title.y = element_text(color = "black", size = 18, face = "bold"),
        legend.text=element_text(size=16),
        legend.title = element_text(size=18),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank())

t<-ggplot(pulse_new,aes(x=week))
t<-t+geom_line(aes(y=an,color="Anxiety"),size=1)+geom_line(aes(y=de,color="Depression"),size=1)+geom_line(aes(y=Meds,color="Taking Medication"),size=1)+geom_line(aes(y=Therapy,color="Receiving Therapy"),size=1)+geom_line(aes(y=Unmet,color="Unmet Need"),size=1)+geom_vline(xintercept=as.Date("2020-03-07"), linetype="dashed",size=1)+xlab("Time")+ylab("\n Percent Reporting")+scale_y_continuous(labels = scales::percent,breaks = c(0.1,0.15,0.2,0.25,0.3,0.35,0.4))+scale_x_date(date_labels = "%b",breaks = as.Date(c("2020-01-07","2020-02-07","2020-03-07","2020-04-07","2020-05-07","2020-06-07","2020-07-07","2020-08-07","2020-09-07","2020-10-07","2020-11-07","2020-12-07")))+scale_color_discrete(name="Pulse Survey Item",labels=c("Anxiety","Depression","Taking Medication","Receiving Therapy","Unmet Need"))+theme_bw()+ 
  theme(axis.text.x = element_text(color = "grey20", size =16),
        axis.text.y = element_text(color = "grey20", size = 16),  
        axis.title.x = element_text(color = "grey20", size = 16),
        axis.title.y = element_text(color = "black", size = 18,face = "bold"),
        legend.text=element_text(size=16),
        legend.title = element_text(size=18),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank())

ggarrange(r, s, t,
          labels = c("A", "B","C"),
          ncol = 1, nrow = 3, align="v")
ggsave("figure2_new.eps",height=12,width = 15)
```

Make overall dataset for SSD, NSSP and pulse survey
```{r}
table<-merge(pulse,ssd_new,by.x = "week", by.y = "week",all=TRUE)
table<-merge(table,nssp_new,by.x = "week",by.y = "week",all=TRUE)
table<-table[,-11]
table<-table[,-13:-19]
table<-table %>% rename("Pulse Anxiety"=an,
                        "Pulse Depression"=de,
                        "Pulse Medication"=Meds,
                        "Pulse Therapy"=Therapy,
                        "Pulse Needed Treatment"=Unmet,
                        "SSD Anxiety"=symptom.Anxiety,
                        "SSD Depression"=symptom.Depression,
                        "SSD Suicidal Ideation"=symptom.Suicidal.ideation,
                        "SSD Motion Sickness"=symptom.Motion.sickness,
                        "NSSP Disaster related mental health ED visits"=Disaster.Related.Mental.Health.Age.10..Count,
                        "NSSP Suicidal attempt ED visits"=Suicide.Attempt.Age.10..Count)
write_csv(table, "SSD_Pulse_NSSP.csv")
```

#lag effect
```{r}
rm(list=ls())
whole<-read.csv("SSD_Pulse_NSSP.csv")
whole<-whole %>% mutate(lag1_an=lag(SSD.Anxiety,1),lag2_an=lag(SSD.Anxiety,2),
                        lag1_de=lag(SSD.Depression,1),lag2_de=lag(SSD.Depression,2),
                        lag1_su=lag(SSD.Suicidal.Ideation,1),lag2_su=lag(SSD.Suicidal.Ideation,2))

#ssd anxiety
model_mh<-glm(NSSP.Disaster.related.mental.health.ED.visits~SSD.Anxiety+lag1_an+lag2_an, data=whole, family = gaussian(link = "log"))
summary(model_mh)
exp(model_mh$coefficients)
exp(confint(model_mh))
model_sa<-glm(NSSP.Suicidal.attempt.ED.visits~SSD.Anxiety+lag1_an+lag2_an, data=whole, family = gaussian(link = "log"))
summary(model_sa)
exp(model_sa$coefficients)
exp(confint(model_sa))

#ssd depression
model_mh<-glm(NSSP.Disaster.related.mental.health.ED.visits~SSD.Depression+lag1_de+lag2_de, data=whole, family = gaussian(link = "log"))
summary(model_mh)
exp(model_mh$coefficients)
exp(confint(model_mh))
model_sa<-glm(NSSP.Suicidal.attempt.ED.visits~SSD.Depression+lag1_de+lag2_de, data=whole, family = gaussian(link = "log"))
summary(model_sa)
exp(model_sa$coefficients)
exp(confint(model_sa))

#ssd sucidal ideation
model_mh<-glm(NSSP.Disaster.related.mental.health.ED.visits~SSD.Suicidal.Ideation+lag1_su+lag2_su, data=whole, family = gaussian(link = "log"))
summary(model_mh)
exp(model_mh$coefficients)
exp(confint(model_mh))
model_sa<-glm(NSSP.Suicidal.attempt.ED.visits~SSD.Suicidal.Ideation+lag1_su+lag2_su, data=whole, family = gaussian(link = "log"))
summary(model_sa)
exp(model_sa$coefficients)
exp(confint(model_sa))
```
#prepare correlation
```{r}
library(GGally)
rm(list=ls())
table<-read.csv("SSD_Pulse_NSSP.csv")
cor_df<-table %>% select(-c("week"))
cor_df<-cor_df %>% rename("HPS Anxiety"=Pulse.Anxiety,
                        "HPS Depression"=Pulse.Depression,
                        "HPS Taking Medication"=Pulse.Medication,
                        "HPS Receiving Therapy"=Pulse.Therapy,
                        "HPS Unmet Need"=Pulse.Needed.Treatment,
                        "SSD Anxiety"=SSD.Anxiety,
                        "SSD Depression"=SSD.Depression,
                        "SSD Suicidal Ideation"=SSD.Suicidal.Ideation,
                        "SSD Motion Sickness"=SSD.Motion.Sickness,
                        "NSSP Disaster related mental health ED visits"=NSSP.Disaster.related.mental.health.ED.visits,
                        "NSSP Suicidal attempt ED visits"=NSSP.Suicidal.attempt.ED.visits)
cor_df<-cor_df[,c(11,10,5,4,3,2,1,8,9,7,6)]
```

#triangle correlation matrix with ggheatmap
```{r}
#ref:http://www.sthda.com/english/wiki/ggplot2-quick-correlation-matrix-heatmap-r-software-and-data-visualization
#ref: https://rpkgs.datanovia.com/rstatix/reference/cor_mat.html
cormat <- round(cor(cor_df,use="pairwise.complete.obs", method = "pearson"),2)
head(cormat)
library(reshape2)
melted_cormat <- melt(cormat)
head(melted_cormat)

# Get lower triangle of the correlation matrix
  get_lower_tri<-function(cormat){
    cormat[upper.tri(cormat)] <- NA
    return(cormat)
  }
  # Get upper triangle of the correlation matrix
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }


# Reorder the correlation matrix
lower_tri <- get_lower_tri(cormat)

#get p value
library(rstatix)
cor.mat<-cor_df %>% cor_mat()
rownames(cor.mat)<-NULL
cor.mat<- as.data.frame(cor.mat) %>% column_to_rownames(., var = "rowname")

p_value<-cor.mat %>% cor_get_pval()
rownames(p_value)<-NULL
p_value<- as.data.frame(p_value) %>% column_to_rownames(., var = "rowname")

lower_p<-get_lower_tri(p_value)
test<-as.matrix(lower_p)
list_p <- as.matrix(lower_p)
melted_p <- melt(list_p,na.rm = TRUE)


# Melt the correlation matrix
melted_cormat <- melt(lower_tri, na.rm = TRUE)

#merge correlation and p value
melted_df<-left_join(melted_cormat,melted_p,by=c("Var1","Var2"))
melted_df<-melted_df %>%  rename(value=value.x, p=value.y)
melted_df$p_value<- ifelse(melted_df$p!=0,formatC(round(melted_df$p,2),format = 'f',2),NA)
melted_df <- melted_df %>% mutate(label=case_when(p!=0 ~ paste0(formatC(value,format = 'f',2),
                                " \n(", p_value,')'), TRUE~'1'))
melted_df[31,6]<-'1'

# Create a ggheatmap
ggheatmap <- ggplot(melted_df, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
    name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 16, hjust = 1),
    axis.text.y = element_text(size=16))+
 coord_fixed()

ggheatmap + 
geom_text(aes(Var2, Var1, label = label), color = "black", size = 4) +
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(1.0,0.3),
  legend.direction = "horizontal",
  legend.text=element_text(size=12),
        legend.title = element_text(size=18))+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))
ggsave("figure3_new.pdf",height=12,width = 15)

```


#figure 4 scatterplot
```{r}
rm(list=ls())
table<-read.csv("SSD_Pulse_NSSP.csv")
s1<-ggplot(table, aes(x=SSD.Anxiety, y=NSSP.Suicidal.attempt.ED.visits)) + 
  geom_point()+xlab("Normalized Google Search for Anxiety")+ylab("Number of Suicidal Attempt ED Visits \n in NSSP")+theme_bw()+geom_smooth(method=lm)+
  theme(axis.text.x = element_text(color = "grey20", size =12),
        axis.text.y = element_text(color = "grey20", size = 16),  
        axis.title.x = element_text(color = "grey20", size = 16),
        axis.title.y = element_text(color = "grey20", size = 14))
s2<-ggplot(table, aes(x=SSD.Anxiety, y=NSSP.Disaster.related.mental.health.ED.visits)) + 
  geom_point()+xlab("Normalized Google Search for Anxiety")+ylab("Number of Disaster Related Mental Health \n ED Visits in NSSP")+theme_bw()+geom_smooth(method=lm)+
  theme(axis.text.x = element_text(color = "grey20", size =12),
        axis.text.y = element_text(color = "grey20", size = 16),  
        axis.title.x = element_text(color = "grey20", size = 16),
        axis.title.y = element_text(color = "grey20", size = 14))

s6<-ggplot(table, aes(x=SSD.Depression, y=Pulse.Therapy)) + 
  geom_point()+xlab("Normalized Google Search for Depression")+ylab("Percentage of respondents \n who received mental therapy in Pulse Survey")+scale_y_continuous(labels = scales::percent)+theme_bw()+
  geom_smooth(method=lm)+
  theme(axis.text.x = element_text(color = "grey20", size =12),
        axis.text.y = element_text(color = "grey20", size = 16),  
        axis.title.x = element_text(color = "grey20", size = 16),
        axis.title.y = element_text(color = "grey20", size = 14))

ggarrange(s1, s2,s6, 
          labels = c("A", "B", "C"),
          ncol = 3, nrow = 1)
ggsave("figure4.pdf", width = 18, height = 5)

```
















